---
title: "CKME136-Part 5-Feature selection 4-Random forest-as of 10-03-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
BRFSS29_SmoteV3 <- read.csv(file="C:/Users/user/Desktop/Capstone Project/Submission/R codes/BRFSS29-Smote-V3.csv", header=TRUE, sep=",")
```

```{r}
library(randomForest)
```


# Using random forest for variable selection
```{r}
rfModel_V3 <-randomForest(MICHD ~ ., data = BRFSS29_SmoteV3)
```
# Getting the list of important variables

```{r}
rfModel_V3
```
The higher the value of mean decrease accuracy or mean decrease gini score, the higher the importance of the variable to our model.
```{r}
# Getting the list of important variables
importance(rfModel_V3)
```
Above the median

Region  
AGE        
RELP
EDU       
EMPL   
INCOME    
SMOKESTAT     
SLEPTIM1        
GENHLTH
PHYSHLTHpa30  
DIABETE
PUMOND
STROKE
LASTDENTV
RMVTETH
BMICAT






