---
title: "CKME136-Part 7- Classifier 1- Decision tree"
author: "Minh Trung DANG"
date: "12/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
BRFSS29_Smote <- read.csv(file="C:/Users/user/Desktop/Capstone Project/Submission/R codes/BRFSS29-Smote-V3.csv", header=TRUE, sep=",")
```

```{r}
str(BRFSS29_Smote)
```

```{r}
library(caret)
```

```{r}
Trainindex <- createDataPartition(y = BRFSS29_Smote$MICHD , p = .70, list = FALSE)
training <- BRFSS29_Smote[Trainindex ,]
validation <- BRFSS29_Smote[-Trainindex,]
```

```{r}
training_new <- training[-32]
validation_new <- validation[-32]
```

```{r}
MICHD_training_label <- training$MICHD
MICHD_validation_label <- validation$MICHD
```

```{r}
library(rpart)
library(rpart.plot)
```

Fully grown trees

```{r}
fulltree <- rpart(MICHD ~ AGE + EMPL+ GENHLTH + DIABETE +  PULMOND + STROKE +
                RMVTETH + LASTDENTV, 
                    data = training,
                    method = "class")
```

```{r, fig.height= 7.5}
# Plot the trees
fancyRpartPlot(fulltree, caption = NULL)
```

GENHLTH, EMPL, PULMON, STROKE, DIABETE

```{r}
# Make predictions on the test data
predicted.classes <- fulltree %>% 
  predict(validation, type = "class")
head(predicted.classes)
```
```{r}
# Compute model accuracy rate on test data
mean(predicted.classes == validation$MICHD)
```
The accuracy is 0.75

#Confusion matrix
```{r}
library(caret)
```

```{r}
confusionMatrix(predicted.classes, validation$MICHD)
```

The three looks very simple. Do we need to prune the tree ?

One possible robust strategy of pruning the tree (or stopping the tree to grow) consists of avoiding splitting a partition if the split does not significantly improves the overall quality of the model.
In rpart package, this is controlled by the complexity parameter (cp), which imposes a penalty to the tree for having two many splits. The default value is 0.01. The higher the cp, the smaller the tree.

A too small value of cp (larger tree) leads to overfitting and a too large cp value will result to a too small tree. Both cases decrease the predictive performance of the model.

An optimal cp value can be estimated by testing different cp values and using cross-validation approaches to determine the corresponding prediction accuracy of the model. The best cp is then defined as the one that maximize the cross-validation accuracy 

trControl, to set up 10-fold cross validation
tuneLength, to specify the number of possible cp values to evaluate. Default value is 3, here we’ll use 10.


```{r}
# Fit the model on the training set
set.seed(123)
fulltree2 <- train(MICHD ~ AGE + EMPL+ GENHLTH + DIABETE +  PULMOND + STROKE +
                RMVTETH + LASTDENTV,
                data = training, 
                method = "rpart",
                trControl = trainControl("cv", number = 10),
                tuneLength = 100)

# Plot model accuracy vs different values of
# cp (complexity parameter)
plot(fulltree2)
```


```{r}
# Print the best tuning parameter cp that
# maximizes the model accuracy
fulltree2$bestTune
```









